---
title: "assignment_2"
author: "Amy Gardiner"
date: "07/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Visualisation
```{r, echo=FALSE}
library(tidyverse)
library(Stat2Data)
data("Hawks")
hawksSmall<-drop_na(select(Hawks,Age,Day,Month,Year,CaptureTime,Species,Wing,Weight,Tail))
```

The number of rows and columns within `hawksSmall` are:
```{r, echo=FALSE}
dim(hawksSmall)
```

Some of the first rows of `hawksSmall` are:
```{r, echo=FALSE}
head(hawksSmall)
```

A histogram plot of the wing length data:


```{r, echo=FALSE}
hawkplot<-ggplot(data=hawksSmall,aes(x=Weight))+xlab("Weight(gm)")
hawkplot+geom_histogram(binwidth=100)+ylab("Count")
```

A density plot of the tail length data:

```{r, echo=FALSE}
hawkplot<-ggplot(data=hawksSmall,aes(x=Tail))+xlab("Tail(mm)")
hawkplot+geom_density()+ylab("Density")
```

This data with `adjust=0.5` and `adjust=1` :
(adjust = A multiplicate bandwidth adjustment. This makes it possible to adjust the bandwidth while still using the a bandwidth estimator. For example, adjust = 1/2 means use half of the default bandwidth.)

```{r, echo=FALSE}
hawkplot<-ggplot(data=hawksSmall,aes(x=Tail))+xlab("Tail(mm)")
hawkplot+geom_density(adjust=0.5)+ylab("Density")

hawkplot<-ggplot(data=hawksSmall,aes(x=Tail))+xlab("Tail(mm)")
hawkplot+geom_density(adjust=1)+ylab("Density")
```

Tail lengths between species in density and violin form:

```{r, echo=FALSE}
ggplot(data=hawksSmall,aes(x=Tail,color=Species))+geom_density()+xlab("Tail(mm)")+ylab("Density")

```

Presenting this data in scatter form:

```{r, echo=FALSE}
tails_scatter<-ggplot(data=hawksSmall,aes(y=Weight,x=Tail))+xlab("Tail(mm)")+ylab("Weight(gm)")
tails_scatter+geom_point(aes(color=Species,shape=Species))
```

...and separating into species-specific facets:

```{r, echo=FALSE}
tails_scatter+geom_point(aes(color=Species))+geom_smooth(method="lm")+facet_wrap(~Species)
```

## Data Wrangling

hSF data frames:


```{r, echo=FALSE}
hSF<-select(filter(Hawks,Species=="RT",Weight>=1000),Wing,Weight,Tail)
head(hSF)
hSF<-arrange(hSF,Wing)
head(hSF)
```

Joining data frames:

```{r, echo=FALSE}
species<-unique(Hawks$Species)
species_name<-c("Red-tailed","Cooper's","Sharp-shinned")
species_names_df<-data.frame(species,species_name)
species_names_df
capitalised<-rename(species_names_df,Species=species)
hawksFullName<-select(left_join(capitalised,Hawks, by="Species"),-Species)
head(hawksFullName)
head(select(hawksFullName,species_name,Wing,Weight))
```

The mutate function

```{r,echo=FALSE}
hawksWithBMI <- Hawks %>%
  mutate(bird_BMI=1000*Weight/Wing^2) %>%
  select(Species,bird_BMI) %>%
  arrange(desc(bird_BMI))

hawksWithBMI %>% head(8)
hawksWithBMI <- hawksWithBMI %>%
  filter(bird_BMI <100)

ggplot(data=hawksWithBMI,aes(x=bird_BMI,y=Species,fill=Species))+geom_violin()+xlab("Bird BMI")+scale_x_continuous(trans='log2')
```

Summarize and group-by functions

```{r, echo=FALSE}
hawksFullName %>%
  group_by(species_name) %>%
  summarize(num_rows=n(),mn_wing=mean(Wing,na.rm=1),md_wing=median(Wing,na.rm=1),t_mn_wing=mean(Wing,na.rm=1,trim=0.1),tail_wing_ratio=mean(Wing/Tail,na.rm=1))

hawksFullName %>%
  select(species_name,Wing,Weight,Culmen,Hallux,Tail,StandardTail,Tarsus,Crop) %>%
  group_by(species_name) %>%
  summarize(across(everything(),~sum(is.na(.x)))) %>%
  head()
```

## Exploratory data analysis

```{r, echo=FALSE}
Hawks %>%
  group_by(Species) %>%
  summarise(across(c("Wing","Weight"), list(mean=~mean(.x,na.rm=1), t_mean=~mean(.x,na.rm=1,trim=0.1), med=~median(.x,na.rm=1))))
```

Robustness of location estimators

```{r, echo=FALSE}
hal<-Hawks$Hallux
hal<-hal[!is.na(hal)]
outlier_val<-100
num_outliers<-10
corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
mean(hal)
mean(corrupted_hal)

num_outliers_vect<-seq(0,1000)
means_vect<-c()
medians_vect<-c()
t_means_vect<-c()
for(num_outliers in num_outliers_vect){
  corrupted_hal<-c(hal,rep(outlier_val,times=num_outliers))
  means_vect<-c(means_vect,mean(corrupted_hal))
  medians_vect<-c(medians_vect,median(corrupted_hal))
  t_means_vect<-c(t_means_vect,mean(corrupted_hal,trim=0.1))
}

df_means_medians<-data.frame(num_outliers=num_outliers_vect,mean=means_vect,t_mean=t_means_vect,median=medians_vect)

df_means_medians %>%
  pivot_longer(!num_outliers,names_to="Estimator", values_to="Value") %>%
  ggplot(aes(x=num_outliers,color=Estimator,linetype=Estimator,y=Value)) + geom_line() + xlab("Number of outliers")
```

Box plots and outliers

```{r,echo=FALSE}
hawk_weights<-select(Hawks,Species,Weight)
ggplot(data=hawk_weights,aes(x=Species,y=Weight))+geom_boxplot()+xlab("Species")+ylab("Weight")

outliers<-function(x){
  q25<-quantile(x,0.25,na.rm=1)
  q75<-quantile(x,0.75,na.rm=1)
  iq_range<-q75-q25
  return(x[((x>q75+1.5*iq_range)|(x<q25-1.5*iq_range))&!is.na(x)])
}

num_outliers<-function(x){
  return(length(outliers(x)))
}

Hawks %>%
  group_by(Species) %>%
  summarise(num_outliers_weight=num_outliers(Weight))
```


