---
title: "assignment_2"
author: "Amy Gardiner"
date: "07/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, echo=FALSE}
library(tidyverse)
library(Stat2Data)
data("Hawks")
hawksSmall<-drop_na(select(Hawks,Age,Day,Month,Year,CaptureTime,Species,Wing,Weight,Tail))
```

The number of rows and columns within `hawksSmall` are:
```{r, echo=FALSE}
dim(hawksSmall)
```

Some of the first rows of `hawksSmall` are:
```{r, echo=FALSE}
head(hawksSmall)
```

A histogram plot of the wing length data:


```{r, echo=FALSE}
hawkplot<-ggplot(data=hawksSmall,aes(x=Weight))+xlab("Weight(gm)")
hawkplot+geom_histogram(binwidth=100)+ylab("Count")
```

A density plot of the tail length data:

```{r, echo=FALSE}
hawkplot<-ggplot(data=hawksSmall,aes(x=Tail))+xlab("Tail(mm)")
hawkplot+geom_density()+ylab("Density")
```

This data with `adjust=0.5` and `adjust=1` :
(adjust = A multiplicate bandwidth adjustment. This makes it possible to adjust the bandwidth while still using the a bandwidth estimator. For example, adjust = 1/2 means use half of the default bandwidth.)

```{r, echo=FALSE}
hawkplot<-ggplot(data=hawksSmall,aes(x=Tail))+xlab("Tail(mm)")
hawkplot+geom_density(adjust=0.5)+ylab("Density")

hawkplot<-ggplot(data=hawksSmall,aes(x=Tail))+xlab("Tail(mm)")
hawkplot+geom_density(adjust=1)+ylab("Density")
```

Tail lengths between species in density and violin form:

```{r, echo=FALSE}
ggplot(data=hawksSmall,aes(x=Tail,color=Species))+geom_density()+xlab("Tail(mm)")+ylab("Density")

ggplot(data=hawksSmall,aes(x=Tail,y=Species,fill=Species))+geom_violin()+xlab("Tail(mm)")
```
