---
title: "Assignment 7"
author: "Amy Gardiner"
date: "18/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(Stat2Data)
library(palmerpenguins)
```

## 1. Student's t-confidence intervals

Importing and filtering for the weights of the red-tailed hawks
```{r}
data("Hawks")
Hawks<-Hawks%>%
  filter(Species=="RT")
RT_Weights<-Hawks%>%
  pull(Weight)

RT_Weights<-RT_Weights[!is.na(RT_Weights)]

```

Student's t method with 99%-level confidence intervals:
```{r}
alpha<-0.01
sample_size<-length(RT_Weights)
sample_mean<-mean(RT_Weights)
sample_sd<-sd(RT_Weights)
t<-qt(1-alpha/2,df=sample_size-1) #df here is degrees of freedom
confidence_interval_l<-sample_mean-t*sample_sd/sqrt(sample_size)
confidence_interval_u<-sample_mean+t*sample_sd/sqrt(sample_size)
confidence_interval<-c(confidence_interval_l,confidence_interval_u)
confidence_interval
```

Comparing assumptions of this statistical test (Gaussian) with a density and quantile-quantile plot:
```{r}
tibble(RT_Weights)%>%ggplot(aes(x=RT_Weights))+geom_density()+theme_bw()+labs(x="Weight",y="Density")

tibble(RT_Weights)%>%ggplot(aes(sample=RT_Weights))+stat_qq()+stat_qq_line(color="blue")+theme_bw()
```

## 2. One sample t-test

```{r}
data("penguins")
head(penguins)

penguins<-penguins%>%
  filter(species=="Adelie")
bill_adelie<-penguins%>%
  pull(bill_length_mm)

bill_adelie<-bill_adelie[!is.na(bill_adelie)]

t.test(x=bill_adelie,mu=40,conf.level=0.99)
```

## 3. Implementing a one-sample t-test

```{r}
my_t_test<-function(vect,mean){
  sample_size<-length(vect)
  sample_mean<-mean(vect)
  sample_sd<-sd(vect)
  test_statistic<-(sample_mean-mean)/(sample_sd/sqrt(sample_size))
  p_value<-2*(1-pt(abs(test_statistic),df=sample_size-1))
  results<-c(test_statistic,p_value)
  results
}

my_t_test(bill_adelie,40)
```

This implementation of the previous example confirms the validity of my function

## The paired t-test

```{r}
library(PairedData)
data("Barley")
head(Barley)
Glabron<-pull(Barley, Glabron)
Velvet<-pull(Barley, Velvet)

t.test(Glabron, Velvet, paired= TRUE,conf.level=0.99)

diffs<-c(Glabron-Velvet)
y_bar<-mean(diffs)
s<-sd(diffs)
effect_size<-y_bar/s
effect_size
```


