---
title: "assignment_5"
author: "Amy Gardiner"
date: "26/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) 
library(ggplot2)
library(tidyverse)
```

## Expectation and variance of a discrete random variable

Suppose that α,β ∈ [0,1] with α+β ≤ 1 and let X be a discrete random variable with with distribution supported on {0, 1, 5}. Suppose that P (X = 1) = α and P (X = 5) = β and P (X ∈/ {0, 1, 5}) = 0.
What is the probability mass function pX : S → [0,1] for X?

pX = α if X=1\
   = β if X=5\
   = 1-α-β if X=0\
   = 0 if X ∈/ {0, 1, 5}
   
The expectation is α + 5β

The variance is sum(pX.(x-(α + 5β))^2) = β(5-(α + 5β))^2 + α (1-(α + 5β))^2) (can expand if necessary)

## Simulating data with the uniform distribution

```{r}
set.seed(0)
n<-1000
sample_X<-data.frame(U=runif(n))%>%
  mutate(X=case_when(
    (0<=U)&(U<0.25)~1,
    (0.25<=U)&(U<0.5)~5,
    (0.5<=U)&(U<=1)~0))%>%
pull(X)

head(sample_X,400)

sample_X_015<-function(alpha,beta,n){
  data.frame(U=runif(n))%>%
  mutate(X=case_when(
    (0<=U)&(U<alpha)~1,
    (alpha<=U)&(U<alpha+beta)~5,
    (alpha+beta<=U)&(U<=1)~0))%>%
 # mutate(sample_avg=map_dbl(.x=U,~mean(.x))) %>%
  pull(X) 
}

test<-sample_X_015(0.5,0.1,100000)

head(test,400)

```


