---
title: "Assignment 4"
author: "Amy Gardiner"
date: "18/10/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) 
library(ggplot2)
library(tidyverse)
```

## 1. Bayes Theorem 

Let A be the event that it rains next week and B the event that the weather forecaster predicts that there will be rain next week. Let's suppose that the probability of rain next week is P(A)=0.9. Suppose also that the conditional probability that there is a forecast of rain, given that it really does rain, is P(B|A)=0.8. On the other hand, the conditional probability that there is a forecast of dry weather, given that there really isn't any rain is P(B_c|A_c)=0.75. 

Now suppose that there is a forecast of rain. What is the conditional probability of rain, given the forecast of rain, P(A|B)?

- P(A)=0.9
- P(B|A)=0.8
- P(B_c|A_c)=0.75

First compute P(B) = P(B^A) + P(B^A_c) 
                   = P(B|A) . P(A) + P(B|A_c) . P(A_c)
                   = P(B|A) . P(A) + (1-P(B_c|A_c) ) . (1-P(A))
                   = 0.8 . 0.9 + (1-0.75) . (1-0.9) = 0.745

## Conditional Probabilities

1. P(A|B)=P(A^B)/P(B)

## Sampling with replacement
```{r}
choose(8,3)
```

n balls per bag\
r are red balls\
(n-r) are blue balls\
process of sampling with replacement repeated k times\
want to know probability of z balls being red\
A_z,k = event of z of k balls being red\

P(A_z,k)= binomial(k z) x (r/n)^z x (1-(r/n))^(k-z)

```{r}
prob_red_spheres<-function(z){
  n<-10
  r<-3
  k<-35
  probability<-(choose(k,z)*((r/n)^z)*(1-(r/n))^(k-z))
  return(probability)
}

prob_red_spheres(20)

num_reds<-seq(1,35,1)
prob<-prob_red_spheres(num_reds)

prob_by_num_reds<-data.frame(num_reds,prob)

prob_by_num_reds%>%head(3)

ggplot(data=prob_by_num_reds)+xlab("Number of reds")+geom_line(aes(x=num_reds,y=prob))+ylab("Probability")

sample(10,35,replace=TRUE)

# Setting the random seed just once
set.seed(0)

for(i in 1:5){
  
  print(sample(100,5,replace=FALSE))
  # The result may well differ every time
}

# Resetting the random seed every time

for(i in 1:5){
  set.seed(1)
  print(sample(100,5,replace=FALSE))
}
```

Simulation study using `sample()`
```{r}
num_trials<-1000 # set the number of trials
set.seed(0) # set the random seed

sampling_with_replacement_simulation<-data.frame(trial=1:num_trials)%>%
  mutate(sample_balls=map(.x=trial,~sample(10,35,replace=TRUE)))
# generate collection of num_trials simulations

sampling_with_replacement_simulation<-sampling_with_replacement_simulation%>%
  mutate(num_reds=map_dbl(.x=sample_balls,~sum(.x<=3)))

num_reds_in_simulation<-sampling_with_replacement_simulation%>%
  pull(num_reds)
# extracts a vector corresponding to the number of reds in each trial

prob_by_num_reds<-prob_by_num_reds%>%
  mutate(simulation_count=map_dbl(.x=num_reds,~sum(num_reds_in_simulation==.x)))
# adds a column which gives the number of trials with a given number of reds

prob_by_num_reds<-prob_by_num_reds%>%
  mutate(expected_count=num_trials*prob)
# adds a column which gives the expected number of reds
# based on the probability formula

prob_by_num_reds%>%
  rename(Simulation=simulation_count,Expected=expected_count)%>%
  pivot_longer(cols=c("Simulation","Expected"),names_to="Type",values_to="count") %>%
  ggplot(aes(num_reds,count)) + geom_line(aes(linetype=Type, color=Type)) + scale_linetype_manual(values = c("solid", "dashed")) + theme_bw() + xlab("Number of reds") + ylab("Count")
```

## Sampling without replacement

```{r}
set.seed(0)
number_trials<-25
number_reds<-50
number_blues<-30
number_greens<-20

sampling_without_replacement_simulation<-data.frame(trial=1:number_trials)%>%
  mutate(sample_balls=map(.x=trial,~sample(100,10,replace=FALSE)))

sampling_without_replacement_simulation<-sampling_without_replacement_simulation%>%
  mutate(number_reds=map_dbl(.x=sample_balls,~sum(.x<=50))) %>%
  mutate(number_blues=map_dbl(.x=sample_balls,~sum(.x>50&&.x<=80))) %>%
  mutate(number_greens=map_dbl(.x=sample_balls,~sum(.x>80&&.x<=100))) %>%
  mutate(minimum_count=pmin(number_reds,number_blues,number_greens)) %>%
  mutate(proportion_missing=map_dbl(.x=minimum_count,~sum(.x==0)))

sampling_without_replacement_simulation

```



